---
title: "R4MS WS2 - Final Exercise"
author: "Jessica Riker"
format: html
editor: visual
---

# Introduction

2.12 Keystone exercise: Estuary Fish Survey - Data Rescue Mission
The Premise: You have inherited a project from a former PhD student who was studying the relationship between water quality and predatory fish assemblages along the Ross River Estuary gradient. Unfortunately, their data management was entirely manual, and the files are a disaster.


```{r}
# Load packages
library(tidyverse) 
library(readxl) 
```

#Phase 1: Ingestion and Decontamination
```{r}
#look at the files in the data folder
#The catch data is split across 4 tabs in one excel file Load indivudally
catch1 <- read_excel(here::here("data/workshop2/estuary_catch_log.xlsx"), sheet = "Lower_ross")
catch2 <- read_excel(here::here("data/workshop2/estuary_catch_log.xlsx"), sheet = 2)
catch3 <- read_excel(here::here("data/workshop2/estuary_catch_log.xlsx"), sheet = 3)
catch4 <- read_excel(here::here("data/workshop2/estuary_catch_log.xlsx"), sheet = 4)
```

```{r}
# Inspect individual tabs
catch1
catch2
catch3
catch4
```

```{r}
#bind rows
# Clean up the environment
catch_data <- bind_rows(catch1, catch2, catch3, catch4)
# Clean up the environment
rm(catch1, catch2, catch3, catch4) 
```

```{r}
#check data
catch_data
head(catch_data)
glimpse(catch_data)
str(catch_data)
```

```{r}
#look at the distinct values
catch_data |> 
  distinct(site)

catch_data |> 
  distinct(species)
```

```{r}
#load metadata
metadata <- read_csv(here::here("data/workshop2/estuary_metadata.csv"))

#check metadata
head(metadata)
glimpse(metadata)
```

```{r}
#Fix site names and convert character to factor variables
catch_data <- catch_data |>
  mutate(
    site = str_to_lower(str_trim(site)),
    site = str_replace_all(site, " ", "_")) |> 
  # Convert all chr to fct
  mutate(across(where(is.character), as_factor)
  )

metadata <- metadata |>
  mutate(
    site_name = str_to_lower(str_trim(site_name)),
    site_name = str_replace_all(site_name, " ", "_")) |> 
  # Convert all chr to fct
  mutate(across(where(is.character), as_factor)
  )

#run a check
catch_data |> 
  str()

str(metadata)
```

```{r}
#fix names by loading dictionary and check
species_dict <- read_csv(here::here("data/workshop2/species_dictionary.csv"))

head(species_dict)
```

```{r}
#look at names and compare
catch_data |> 
  distinct(species)

species_dict |>
  distinct(common_name)

#fix names
catch_data <- catch_data |>
  mutate(
    species = str_to_lower(str_trim(species)),
    species = str_replace_all(species, " ", "_")
  )

#join names
catch_data <- catch_data |>
  mutate(common_name = species) |>
  left_join(species_dict, by = join_by(common_name))

#check
glimpse(catch_data)
```

```{r}
#load in other data
source(here::here("code/diw_sensor.R"))
```


```{r}
#join catch with sensor data and meta data
catch_data_final <- catch_data |>
  left_join(metadata, by = join_by(site == site_name)) |>
  left_join(sensor_daily, by = join_by(site, date))

#check
glimpse(catch_data_final)

#Should have more rows? 480 not 427... missing values or 0s?
```


```{r}
# fix missing values in catch data
catch_data_final <- 
catch_data_final |> 
  complete(
    nesting(site, date, 
           # lat, lon, 
            zone, mean_temp, mean_salinity, mean_turbidity),
    scientific_name
  ) |>
  mutate(
    # Convert the resulting NA counts to 0
    count = coalesce(count, 0)
  )

#check
glimpse(catch_data_final)
#rows now at 480 :)
```


#salinity
```{r}
# Make a table
sal_sum <- 
  catch_data_final |>
  group_by(zone) |>
  summarise(
    sd_salinity = sd(mean_salinity, na.rm = TRUE),
    median_salinity = median(mean_salinity, na.rm = TRUE),
    .groups = "drop"
  )

# Make a plot
sal_sum |> 
  ggplot(aes(x = zone, y = median_salinity)) +
  geom_point() +
  geom_errorbar(aes(ymin = median_salinity - sd_salinity, ymax = median_salinity + sd_salinity), width = 0.2) +
  theme_classic() +
  labs(
       x = "Zone",
       y = "Median Salinity (ppt)")
```

#Turbidity
```{r}
turb_sum <- 
  catch_data_final |>
  group_by(zone) |>
  summarise(
    sd_turbidity = sd(mean_turbidity, na.rm = TRUE),
    median_turbidity = median(mean_turbidity, na.rm = TRUE),
    .groups = "drop"
  )

turb_sum |>
  ggplot(aes(x = zone, y = median_turbidity)) +
  geom_point() +
  geom_errorbar(aes(ymin = median_turbidity - sd_turbidity, ymax = median_turbidity + sd_turbidity), width = 0.2) +
  theme_minimal() +
  labs(
       x = "Zone",
       y = "Median Turbidity (NTU)")
```

#temp
```{r}
temp_sum <- 
  catch_data_final |>
  group_by(zone) |>
  summarise(
    sd_temperature = sd(mean_temp, na.rm = TRUE),
    median_temperature = median(mean_temp, na.rm = TRUE),
    .groups = "drop"
  )

temp_sum |>
  ggplot(aes(x = zone, y = median_temperature)) +
  geom_point() +
  geom_errorbar(aes(ymin = median_temperature - sd_temperature, ymax = median_temperature + sd_temperature), width = 0.2) +
  theme_minimal() +
  labs(
       x = "Zone",
       y = "Median Temperature (°C)")
```
#only salinity didnt have the error bars cross each other so this is what we are looking at

```{r}
# look at stats
fish_stats <- catch_data_final |>
  group_by(zone, scientific_name) |>
  summarise(
    mean_catch = mean(count),
    sem_catch = sd(count) / sqrt(n()),
    .groups = "drop"
  )

# plot
fish_stats |> 
  ggplot(aes(x = zone, y = mean_catch, fill = scientific_name)) +
  geom_line(group = 1) +
  geom_point(data = catch_data_final, aes(x = zone, y = count),
             position = position_jitter(width = 0.2), alpha = 0.2, size = 1, inherit.aes = FALSE) + 
  geom_point(position = "dodge", size = 4, pch = 23) +
  geom_errorbar(aes(ymin = mean_catch - sem_catch, ymax = mean_catch + sem_catch), 
                position = position_dodge(0.9), width = 0.2) + 
  facet_wrap(~ scientific_name, scales = "fixed") +
  theme_minimal() +
  theme(
    strip.text = element_text(face = "italic", size = 10, colour = "black"), 
    legend.position = "none" 
  ) +
  labs(
    x = "Spatial Zone",
    y = expression("Abundance (n day"^-1*")"),
    y = "Abundance (per day)"
  )
```

